\subsection{Standard System Functions}
\texttt{{\large\textbf{to\_int}}}\textsf{(any):integer}\\
 Converts the argument to an integer value.

\hrulefill

\texttt{{\large\textbf{to\_real}}}\textsf{(any):real}\\
 Converts the argument to a real number (double precision).

\hrulefill

\texttt{{\large\textbf{to\_bool}}}\textsf{(any):boolean}\\
 Converts the argument to a boolean value (1=true, 0=false).

\hrulefill

\texttt{{\large\textbf{to\_string}}}\textsf{([any]):string}\\
 Converts the argument[s] to a text string.

\hrulefill

\texttt{{\large\textbf{alloc}}}\textsf{(integer, \{integer\}):array}\\
 Allocates an array of one or two dimensions.

\hrulefill

\texttt{{\large\textbf{path\_only}}}\textsf{(string):string}\\
 Returns the path portion of a complete file path.

\hrulefill

\texttt{{\large\textbf{file\_only}}}\textsf{(string):string}\\
 Returns the file name portion of a complete file path, including extension.

\hrulefill

\texttt{{\large\textbf{ext\_only}}}\textsf{(string):string}\\
 Returns the extension of a file name.

\hrulefill

\texttt{{\large\textbf{extensions}}}\textsf{(none):table}\\
 Returns information about currently loaded extensions

\hrulefill

\texttt{{\large\textbf{ostype}}}\textsf{(none):string}\\
 Returns identifying information about the operating system type. ('osx', 'win32', 'linux', etc).

\hrulefill

\texttt{{\large\textbf{stable\_sort}}}\textsf{(array, [integer:sort column index]):none}\\
 Sort an array of numbers or strings in place while preserving relative ordering of elements.

\hrulefill

\texttt{{\large\textbf{json\_write}}}\textsf{(variant):string}\\
 Convert a variable to a JSON string representation.

\hrulefill

\texttt{{\large\textbf{json\_read}}}\textsf{(string):variant}\\
 Convert a JSON string to a variable.
